---
title: "basics_of_rerddap"
author: "Tylar Murray"
date: "2023-01-11"
output: html_document
---

This document was created as we walked through basic usage of R, Rstudio, ERDDAP, and the rerddap package.
A similar document for working with ERDDAP data in python can be viewed and run using google colab [here](https://colab.research.google.com/drive/1D--oaQNfCw8Qc-fCkR8ps9_ihEtKpXZm?usp=sharing).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# whatever is in this chunk gets run before any other chunk
# put your libary loading here:
#library("rerddap")
```

```{r libraries setup}
install.package("rerddap")
```

```{r check rerrdap is installed by printing version}
rerddap::version()
rerddap::servers()  # list of "default" servers

# to select a single ERDDAP server to use:
database <- "https://upwell.pfeg.noaa.gov/erddap/"
Sys.setenv(RERDDAP_DEFAULT_URL = database)
# this is required when using servers not listed by rerddap::servers()
```

Datasets relevant to IMaRS's Climate Indicators work are listed in [this gSheet](https://docs.google.com/spreadsheets/d/10vbzAxnexaF9vXJckQh03fNHWzsFaHNt9YkUmJAfnKo/edit#gid=0).

The dataset used below can also be viewed in the ERDDAP GUI [here](https://pae-paha.pacioos.hawaii.edu/erddap/griddap/dhw_5km.graph?CRW_SSTANOMALY%5B(2012-05-01T12:00:00Z)%5D%5B(42.025):(-13.975)%5D%5B(-23.975):(1.975)%5D&.draw=surface&.vars=longitude%7Clatitude%7CCRW_SSTANOMALY&.colorBar=%7C%7C%7C%7C%7C&.bgColor=0xffccccff).

```{r rerddap griddap usage}
LAT <- c(22., 51.)
LON <- c(-140., -105)
TIME <- c("2017-01-01", "2017-01-02")
DATASET_ID <- "dhw_5km"

# dhw_5km is a temperature dataset recommended by Dan
dataset_result <- rerddap::griddap(
    DATASET_ID,
    latitude = LAT, 
    longitude = LON, 
    time = TIME, 
    fields = 'CRW_SSTANOMALY',
    fmt = "csv"  # (optional) tells to download as .csv instead of .nc
)
```

```{r save downloaded file to a location}
write.csv(
    dataset_result,
    paste0(DATASET_ID, "_query.csv")
)
```

view the vignette in `??rerddap` "rerddap::Using_rerddap" for next steps.